---
title: "ISLR Chapter 3 with Tidymodel"
output: html_notebook
---

The linear regression Boston housing prices lab in Chapter 3 of ISLR is rewritten here to take advantage of preferred libraries and coding patterns, specifically  [Tidyverse](https://r4ds.had.co.nz/) and [Tidymodels](https://www.tmwr.org/). Also, the [attach](https://www.r-bloggers.com/2011/05/to-attach-or-not-attach-that-is-the-question/) function is best avoided in production code. You can find an R script in my github folder for [ISLR](https://github.com/AndrewBorst/R-Learning/tree/main/ISLR).

We use Boston housing data from the ISLR2 package and predict the median value of homes (medv). 
```{r import-Boston, echo=TRUE}
library(ISLR2)
library(tidyverse)
library(tidymodels)
library(GGally)

?Boston # to see the data dictionary for Boston housing prices data, copied below
glimpse(Boston)
```

Boston

A data frame with 506 rows and 13 variables.

+ crim
  + per capita crime rate by town.
+ zn
  + proportion of residential land zoned for lots over 25,000 sq.ft.
+ indus
  + proportion of non-retail business acres per town.
+ chas
  + Charles River dummy variable (= 1 if tract bounds river; 0 otherwise).
+ nox
  + nitrogen oxides concentration (parts per 10 million).
+ rm
  + average number of rooms per dwelling.
+ age
  + proportion of owner-occupied units built prior to 1940.
+ dis
  + weighted mean of distances to five Boston employment centres.
+ rad
  + index of accessibility to radial highways.
+ tax
  + full-value property-tax rate per $10,000.
+ ptratio
  + pupil-teacher ratio by town.
+ lstat
  + lower status of the population (percent).
+ medv
  + median value of owner-occupied homes in $1000s.

```{r}
Boston %>%  
  select(lstat, medv) %>% 
  ggpairs()
```


End of Lab
---
Notebook hints: 

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
